---
title: "2025_4_3_CodingChallengeSeven_mer0127"
author: "Madeline Redd"
date: "`r format(Sys.time(), '%Y - %m - %d')`"
fontsize: "12pt"
geometry: "margin=1in"
output:
  pdf_document: 
    keep_tex: TRUE
    latex_engine: xelatex
---


## Question One

Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

```{r, warning = FALSE}

library(tidyverse)
library(lme4)
library(emmeans)
library(multcompView)
library(multcomp)

PlantEmerg <- read.csv("PlantEmergence.csv")
str(PlantEmerg)

#Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

PlantEmerg$Treatment <- as.factor(PlantEmerg$Treatment)
PlantEmerg$DaysAfterPlanting <- as.factor(PlantEmerg$DaysAfterPlanting)
PlantEmerg$Rep <- as.factor(PlantEmerg$Rep)

head(PlantEmerg)

```


## Question Two

Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r}

lm_Q2<- lm(Emergence ~ DaysAfterPlanting * Treatment, data = PlantEmerg)

summary(lm_Q2)

aov_Q2<- aov(lm_Q2)

summary(aov_Q2)

```


## Question Three

Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

```{r}
simp_lm_Q3<- lm(Emergence ~ DaysAfterPlanting + Treatment, data = PlantEmerg)

summary(simp_lm_Q3)
```
### Interpertation for Question Three:

Looking at Treatment Two
Intercept: -134.531 
Coefficient:


## Question Four

Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
Q4<-emmeans(simp_lm_Q3, ~ Treatment) 
                    

Results_Q4 <- cld(Q4, alpha = 0.05, reversed = TRUE, details = TRUE) 
                                    #Contrast with Tukey adjustment by default 

Results_Q4
```

## Question Five

The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) #Estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, 
                Letters = letters) # contrast with Tukey adjustment by default.
  
    # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), 
              size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

```



## Question Six

Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.


[Coding Challenge Seven Link](https://github.com/Mads-Hamrick/ENTM6820class/tree/89f9f8cb2ed9120e3835c943c98f6a304e9b0aad/2025_4_3_CodingChallengeSeven_mer0127)


