---
title: "2025_4_3_CodingChallengeSix_mer0127"
author: "Madeline Redd"
date: "`r format(Sys.time(), '%Y - %m - %d')`"
fontsize: "12pt"
geometry: "margin=1in"
output:
  pdf_document: 
    keep_tex: TRUE
    latex_engine: xelatex
  html_document: 
  word_document:
---


## Question One

Regarding reproducibility, what is the main point of writing your own functions and iterations? 

Answer: Creating functions allows others opportunity to see each step of the calculation and will reduce human error if the operations had to be constantly retyped. 
It could benefit someone when reviewing the code and if needed to build on the code. For iterations, it allows replications to be accurate and efficiently if preformed exactly the same, but also allows other to test different parameters.


## Question Two

In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned. 

Answer:

Example of Creating a Function Code -->

new_function <- function(argument_one, argument_two) {
  answer <- argument_one * (argument_two + 87)
                      #Operations to perform with arguments to produce results
  return(answer)  
}

new_function(45, 3)

Example of FOR LOOP Code -->


iteration_length <- lage-fage+1 # Define the iteration length

#Necessary Parameters for the For Loop
#Best for stable coefficients or numbers that might need to be adjusted and help reduce human error

M=0.4 #Natural Mortality
survship0=numeric(iteration_length) #Empty numeric vector for data
survship0[1]=1 #survivorship at age 1

#a is the loop variable 
#the  2:iteration_length: the sequence it will iterate over

 for (a in 2:iteration_length){
  survship0[a]=survship0[a-1]*exp(-M) #Writing calculation code could be built into almost anything.
  
}

print(survship0) #print data



```{r}
#Full code for Mortality with Fishing Pressure on Atlantic Tarpon
#Parameters for for loop
fage=1 #first age
lage=8 #last age
linf=360 #Maximum length 
k=0.45 # growth coefficient
t0=0 #age 0
cv_tl=0.1 #variation in total length
M=0.4 #Natural mortality 
l50_mat=225 #length at 50% pop of maturity
h_mat=0.05 #steepness of maturity 
a_wt=2e-5 #length-weight
b_wt=3 #length-weight
alpha_hat=8 #recruitment
R0=1 #initial recruitment 
tmax=100 #max years
F_full=0.5 #Fishing Mortality
MLL=250 #Minimum legal length 

age=fage:lage
n_ages=lage-fage+1
length=linf*(1-exp(-k*(age-t0)))
weight=a_wt*length^b_wt
maturity=1/(1+exp(-h_mat*(length-l50_mat)))
fecundity=weight*maturity
sd_length=cv_tl*length
vuln=1-pnorm(MLL,length,sd_length)

survship0=numeric(n_ages)
survship0[1]=1

#Survivorship For Loop
for (a in 2:n_ages){
  survship0[a]=survship0[a-1]*exp(-M)
}
survship0[lage]=survship0[lage]/(1-exp(-M))

ssb0=sum(survship0*fecundity)
alpha=alpha_hat/ssb0
beta=log(alpha*ssb0)/(R0*ssb0)


N_a=matrix(NA,nrow=tmax,ncol=n_ages) #Blank matrix for values to be imported into
F_a=N_a
Z_a=N_a
U_a=N_a
yield_t=numeric(tmax)
ssb_t=yield_t
SPR_t=yield_t

N_a[1,]=R0*survship0
F_a[1,]=F_full*vuln
Z_a[1,]=F_a[1,]+M
U_a[1,]=1-exp(-F_a[1,])
yield_t[1]=sum(N_a[1,]*weight*U_a[1,])
ssb_t[1]=sum(N_a[1,]*fecundity)
SPR_t[1]=ssb_t[1]/(R0*ssb0)

for(t in 2:tmax){
  N_a[t,1]=alpha*ssb_t[t-1]*exp(-beta*ssb_t[t-1])
  for(a in 2:n_ages){
    N_a[t,a]=N_a[t-1,a-1]*exp(-Z_a[t-1,a-1])
  }
  N_a[t,n_ages]=N_a[t,n_ages]+N_a[t-1,a]*exp(-Z_a[t-1,a])  
  F_a[t,]=F_full*vuln
  Z_a[t,]=F_a[t,]+M
  U_a[t,]=1-exp(-F_a[t,])
  yield_t[t]=sum(N_a[t,]*weight*U_a[t,])
  ssb_t[t]=sum(N_a[t,]*fecundity)
  SPR_t[t]=ssb_t[t]/(R0*ssb0)
}
  
windows()
plot(SPR_t,type='l',ylim=c(0,1)) 
```




## Question Three

Read in the Cities.csv file from Canvas using a relative file path.

```{r}

CITIES <- read.csv("Cities.csv")
head(CITIES)

```


## Question Four

Write a function to calculate the distance between two pairs of coordinates based on the Haversine formula (see below). The input into the function should be lat1, lon1, lat2, and lon2. The function should return the object distance_km. All the code below needs to go into the function.

```{r}

#GIVEN CODE -->
#Convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

#Haversine Formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

#Earth's radius in kilometers
earth_radius <- 6378137

#Calculate the distance
distance_km <- (earth_radius * c)/1000

```


## Question Five

Using your function, compute the distance between Auburn, AL and New York City 
    a. Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function. 
    b. The output of your function should be 1367.854 km

```{r}

```



## Question Six


Now, use your function within a for loop to calculate the distance between all other cities in the data. The output of the first 9 iterations is shown below. 


```{r}

```



### Bonus Points

Bonus point if you can have the output of each iteration append a new row to a dataframe, generating a new column of data. In other words, the loop should create a dataframe with three columns called city1, city2, and distance_km, as shown below. The first six rows of the dataframe are shown below.





## Question Seven


Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. Provide me a link to your github written as a clickable link in your .pdf or .docx


[Coding Challenge Six]()











